---
title: "R Notebook"
output: html_notebook
---

The top_n verb DplyR

What if instead of aggregatin each state, we wanted to find only the largest county in each state?
DplyR's top n is very useful for keeping the most extreme observations from each group. 

Like summarize(), top_n operates on a grouped table. 

=====================================================================================================
top_n
```{r}
counties_selected <- counties %>%
  select(state, county, population, unemployment, income)

counties_selected %>%
  group_by(state) %>%
  top_n(1, population) #find the county with the highest population in each state.
```
The function takes two arguments: the number observations we want from each group, and the column to weight by.
=====================================================================================================
Highest unemployment
```{r}
counties_selected %>%
  group_by(state) %>%
  top_n(1, unemployment)

#Numbers of observations
counties_selected %>%
  group_by(state) %>%
  top_n(3, unemployment) #< 3 observations for each state
```
 Instead of looking at population, we could get the county with the highest unemployment rate from each state
by changing the second argument to unemployment. 
=====================================================================================================
Selecting a county from each region
```{r}
# Group by region and find the greatest number of citizens who walk to work
counties_selected %>%
  group_by(region)%>%
    top_n(1,walk)
```
# Groups:   region [4]
  region        state        county                 metro    population  walk
  <chr>         <chr>        <chr>                  <chr>         <dbl> <dbl>
1 West          Alaska       Aleutians East Borough Nonmetro       3304  71.2
2 Northeast     New York     New York               Metro       1629507  20.7
3 North Central North Dakota McIntosh               Nonmetro       2759  17.5
4 South         Virginia     Lexington city         Nonmetro       7071  31.7
=====================================================================================================
Finding the highest-income state in each region
```{r}
counties_selected %>%
  group_by(region, state) %>%
  # Calculate average income
  summarize(average_income = mean(income)) %>%
  # Find the highest income state in each region
  top_n(1,average_income)
```
# A tibble: 4 x 3
# Groups:   region [4]
  region        state        average_income
  <chr>         <chr>                 <dbl>
1 North Central North Dakota         55575.
2 Northeast     New Jersey           73014.
3 South         Maryland             69200.
4 West          Alaska               65125.
=====================================================================================================
Using summarize, top_n, and count together
```{r}
# Extract the most populated row for each state
counties_selected %>%
  group_by(state, metro) %>%
  summarize(total_pop = sum(population)) 
```
# A tibble: 97 x 3
# Groups:   state [50]
   state      metro    total_pop
   <chr>      <chr>        <dbl>
 1 Alabama    Metro      3671377
 2 Alabama    Nonmetro   1159243
 3 Alaska     Metro       494990
 4 Alaska     Nonmetro    230471
 5 Arizona    Metro      6295145
 6 Arizona    Nonmetro    346783
 7 Arkansas   Metro      1806867
 8 Arkansas   Nonmetro   1151341
 9 California Metro     37587429
10 California Nonmetro    834035
```{r}
# Extract the most populated row for each state
counties_selected %>%
  group_by(state, metro) %>%
  summarize(total_pop = sum(population)) %>%
  top_n(1, total_pop)
```
# A tibble: 50 x 3
# Groups:   state [50]
   state       metro total_pop
   <chr>       <chr>     <dbl>
 1 Alabama     Metro   3671377
 2 Alaska      Metro    494990
 3 Arizona     Metro   6295145
 4 Arkansas    Metro   1806867
 5 California  Metro  37587429
 6 Colorado    Metro   4590896
 7 Connecticut Metro   3406918
 8 Delaware    Metro    926454
 9 Florida     Metro  18941821
10 Georgia     Metro   8233886
=====================================================================================================
Using Summarize, top_n, and count together
```{r}
# Count the states with more people in Metro or Nonmetro areas
counties_selected %>%
  group_by(state, metro) %>%
  summarize(total_pop = sum(population)) %>%
  top_n(1, total_pop) %>%
  ungroup() %>% 
  count(metro)
```
 A tibble: 2 x 2
  metro        n
  <chr>    <int>
1 Metro       44
2 Nonmetro     6

