---
title: "R Notebook"
output: html_notebook
---
Count() verb: learn to aggregate data, take many observations and summarize them in one.
This is a commom data science strategy for making datasets manageable and interpretable.
It's a very useful type of aggregation when analyzing a dataset.

```{r}
counties %>%
	count(state, sort=TRUE)
```

Add weight: We can add the argument wt, which stands for "weight", equals population.

```{r}
counties %>%
  count(state, wt=population, sort = TRUE)
```

This means that the column after the "state" will be weighted by the population. 
In the result, instead of seeing the number of counties in each state, it shows the total population. 

=========================================================================================================
Counting by region 
```{r}
# Use count to find the number of counties in each region
counties_selected %>%
  count(region, sort = TRUE)
```
# A tibble: 4 x 2
  region            n
  <chr>         <int>
1 South          1420
2 North Central  1054
3 West            447
4 Northeast       217
=========================================================================================================
Counting citizens by state
```{r}
# Find number of counties per state, weighted by citizens
counties_selected %>%
  count(state, wt=citizens, sort = TRUE)
```
# A tibble: 50 x 2
   state                 n
   <chr>             <dbl>
 1 California     24280349
 2 Texas          16864864
 3 Florida        13933052
 4 New York       13531404
 5 Pennsylvania    9710416
 6 Illinois        8979999
 7 Ohio            8709050
 8 Michigan        7380136
 9 North Carolina  7107998
10 Georgia         6978660
# ... with 40 more rows
=========================================================================================================
Mutating and counting: "What are the US states where the most people walk to work?"
```{r}
counties_selected %>%
  # Add population_walk containing the total number of people who walk to work 
  mutate(population_walk = population * walk / 100) %>%
  # Count weighted by the new column
  count(state, wt = population_walk, sort = TRUE)
```
# A tibble: 50 x 2
   state                n
   <chr>            <dbl>
 1 New York      1237938.
 2 California    1017964.
 3 Pennsylvania   505397.
 4 Texas          430783.
 5 Illinois       400346.
 6 Massachusetts  316765.
 7 Florida        284723.
 8 New Jersey     273047.
 9 Ohio           266911.
10 Washington     239764.
# ... with 40 more rows
=========================================================================================================


