---
title: "R Notebook"
output: html_notebook
---

GroupBy, Summarize, Ungroup - Data Manipulation with DplyR

======================================================================================================
Summarize() take many observations and turns them into ONE observation. 
If we wanted to find th total population of the United States, we could use the summarize()
```{r}
counties %>%
  summarize(total_population = sum(population))
```
======================================================================================================
Aggregate and Summarize
It's possible to define multiple variables in a summarize, aggregating each in different ways.  
We can fint the total population, but also the average unemployment rate, using the verb mean()
```{r}
counties %>%
  summarize(total_population = sum(population),
            average_unemployment = mean(unemployment))
```
======================================================================================================
Aggregate within groups
Summarizing the entire table is useful, but, ideally, we want to aggregate within groups, such as to 
find the toal population within each state.
```{r}
counties %>%
    group_by(state) %>%
    summarize(total_pop = sum(population),
              average_unemployment = sum(unemployment))
```
======================================================================================================
Arrange
It's useful to add an additional step of an arrange(), so that we can focus on the MOST NOTABLE examples.
```{r}
counties %>%
  group_by(state)%>%
  summarize(total_pop = sum(population),
            average_unemployment = mean(unemployment)) %>%
  arrange(desc(average_unemployment`))
```
Group By
```{r}
counties %>%
  group_by(state, metro) %>%
  summarize(total_pop = sum(population))
```
Ungroup
If we don't want to keep state as a group, we can add another dplyr verb: ungroup()
```{r}
counties %>%
  group(state, metro) %>%
  summarize(total_pop = sum(population)) %>%
  ungroup()
```
======================================================================================================
Summarizing: useful for collapsing a large dataset into a single observation
```{r}
# Summarize to find minimum population, maximum unemployment, and average income
counties_selected %>%
  summarize(min_population = min(population),
            max_unemployment = max(unemployment),
            average_income = mean(income))
```
======================================================================================================
Summarizing by state
```{r}
# Group by state and find the total area and population
counties_selected %>%
  group_by(state) %>%
  summarize(total_area = sum(land_area),
            total_population = sum(population))
```
# A tibble: 50 x 3
   state       total_area total_population
   <chr>            <dbl>            <dbl>
 1 Alabama         50645.          4830620
 2 Alaska         553560.           725461
 3 Arizona        113594.          6641928
 4 Arkansas        52035.          2958208
 5 California     155779.         38421464
 6 Colorado       103642.          5278906
 7 Connecticut      4842.          3593222
 8 Delaware         1949.           926454
 9 Florida         53625.         19645772
10 Georgia         57514.         10006693
# ... with 40 more rows
```{r}
# Add a density column, then sort in descending order
counties_selected %>%
  group_by(state) %>%
  summarize(total_area = sum(land_area),
            total_population = sum(population)) %>%
  mutate(density = total_population / total_area) %>%
  arrange(desc(density))
```
# A tibble: 50 x 4
   state         total_area total_population density
   <chr>              <dbl>            <dbl>   <dbl>
 1 New Jersey         7354.          8904413   1211.
 2 Rhode Island       1034.          1053661   1019.
 3 Massachusetts      7800.          6705586    860.
 4 Connecticut        4842.          3593222    742.
 5 Maryland           9707.          5930538    611.
 6 Delaware           1949.           926454    475.
 7 New York          47126.         19673174    417.
 8 Florida           53625.         19645772    366.
 9 Pennsylvania      44743.         12779559    286.
10 Ohio              40861.         11575977    283.
# ... with 40 more rows
======================================================================================================
Summarizing by state and region
```{r}
# Summarize to find the total population
counties_selected %>%
  group_by(region, state) %>%
  summarize(total_pop = sum(population)) 
```
# A tibble: 4 x 4
  region        total_pop average_pop median_pop
  <chr>             <dbl>       <dbl>      <dbl>
1 North Central  67532245      64072.     20241 
2 Northeast      55989520     258016.    103855 
3 South         117927776      83048.     26252.
4 West           74395812     166434.     21396
```{r}
# Calculate the average_pop and median_pop columns 
counties_selected %>%
  group_by(region, state) %>%
  summarize(total_pop = sum(population)) %>%
  summarize(average_pop = mean(total_pop),
            median_pop = median(total_pop))
```
# A tibble: 50 x 3
# Groups:   region [4]
   region        state        total_pop
   <chr>         <chr>            <dbl>
 1 North Central Illinois      12873761
 2 North Central Indiana        6568645
 3 North Central Iowa           3093526
 4 North Central Kansas         2892987
 5 North Central Michigan       9900571
 6 North Central Minnesota      5419171
 7 North Central Missouri       6045448
 8 North Central Nebraska       1869365
 9 North Central North Dakota    721640
10 North Central Ohio          11575977
# ... with 40 more rows
