---
title: "R Notebook"
output: html_notebook
---

The transmute() verb:
Transmute is like a combination of select and mutate: getting back just a subset of the columns

- Combination: select & mutate
- Returns a subset of column that are transformed and changed

============================================================================================================================================
Select and calculate:

The transmute() give us the control of what variables we keep,  what NEW variables we calculate, and what to DROP.

We are able to do a calculation at the same time: creating a new column called fraction men. 
This saves somes effort compared to doing both a MUTATE and SELECT.
```{r}
counties %>%
  transmute(state, county, fraction_men = men/population) #Like select, we've dropped all other column and only kept these three. 
```

As another example, we've already worked with the unemployment column that contains the unemployment rate. 
```{r}
counties %>%
  transmute(state, county, population, unemployed_people = population * unemployment / 100)
```
============================================================================================================================================
Using transmute
The transmute verb allows we control which variables we keep, which variables we calculate, and which variables to drop.
```{r}
counties %>%
  # Keep the state, county, and populations columns, and add a density column
  transmute(state, county, population, density = population / land_area) %>%
  # Filter for counties with a population greater than one million 
  filter(population > 1000000) %>%
  # Sort density in ascending order 
  arrange(density)
```
===============================================================================================================================
Choosing among the four verbs
In this chapter you've learned about the four verbs: select, mutate, transmute, and rename. Here, you'll choose the appropriate verb for each situation. You won't need to change anything inside the parentheses.

```{r}
# Change the name of the unemployment column
counties %>%
  rename(unemployment_rate = unemployment)

# Keep the state and county columns, and the columns containing poverty
counties %>%
  select(state, county, contains("poverty"))

# Calculate the fraction_women column without dropping the other columns
counties %>%
  mutate(fraction_women = women / population)

# Keep only the state, county, and employment_rate columns
counties %>%
  transmute(state, county, employment_rate = employed / population)
```
# A tibble: 3,138 x 3
   state   county   employment_rate
   <chr>   <chr>              <dbl>
 1 Alabama Autauga            0.434
 2 Alabama Baldwin            0.441
 3 Alabama Barbour            0.319
 4 Alabama Bibb               0.367
 5 Alabama Blount             0.384
 6 Alabama Bullock            0.362
 7 Alabama Butler             0.384
 8 Alabama Calhoun            0.406
 9 Alabama Chambers           0.402
10 Alabama Cherokee           0.390