---
title: "R Notebook"
output: html_notebook
---
install.packages("dplyr)
install.packages("ggplot2")
library(ggplot2)
library(dplyr) #dplyr verbs are useful for exploring data

Grouped Mutates 
Window Functions

We will reintroduce the ggplot2 package, which we can load with library(ggplot2). 
There are three parts to a ggplot2: the data, the aesthetics, and the layers. 

```{r}
babynames_filtered <- babynames %>%
  filter(name == "Amy")
```

We can call ggplot with this filtered object, and aesthetics x = year and y = babies

```{r}
ggplot(babynames_filtered, aes(x = year, y = babies)) + geom_line()
```
========================================================================================
Filter for multiple names
Besides filtering for just one name, we could filter for multiple names using the % operator
```{r}
babynames_multiple <- babynames %>%
  filter(name %in% c("Leonardo", "Sarah"))
```
========================================================================================
When was each name most common?
We can use the top_n(), along with group_by(), to find the year in which each name was most common
```{r}
babynames %>%
  group_by(name) %>%
  top_n(1, number)
```
========================================================================================
Filtering and arranging for one year
```{r}
babynames %>%
  # Filter for the year 1990
  filter(year == 1990) %>%
  # Sort the number column in descending order 
  arrange(desc(number))
```
========================================================================================
Using top_n with babynames
We could use filter() and arrange() to find the most common names in one year. 
However, we could also use group_by and top_n to find the most common name in every year.
```{r}
# Find the most common name in each year
babynames %>%
  group_by(year) %>%
  top_n(1, number)
```
# Groups:   year [28]
    year name  number
   <dbl> <chr>  <int>
 1  1880 John    9701
 2  1885 Mary    9166
 3  1890 Mary   12113
 4  1895 Mary   13493
 5  1900 Mary   16781
 6  1905 Mary   16135
 7  1910 Mary   22947
 8  1915 Mary   58346
 9  1920 Mary   71175
10  1925 Mary   70857
=======================================================================================
Visualizing names with ggplot2
```{r}
# Filter for the names Steven, Thomas, and Matthew 
selected_names <- babynames %>%
  filter(name %in% c("Steven","Thomas", "Matthew"))
```
A tibble: 84 x 3
    year name    number
   <dbl> <chr>    <int>
 1  1880 Matthew    113
 2  1880 Steven      17
 3  1880 Thomas    2542
 4  1885 Matthew    111
 5  1885 Steven      21
 6  1885 Thomas    2275
 7  1890 Matthew     86
 8  1890 Steven      20
 9  1890 Thomas    2209
10  1895 Matthew    112
# ... with 74 more rows

```{r}
# Plot the names using a different color for each name
ggplot(selected_names, aes(x = year, y = number, color = name)) +
  geom_line()
```

=======================================================================================
Grouped mutates - data manipulation with dplyr

Review: group_by() and summarize()
Tell dplyr to group by the year column then summarize to calculate the sum of the n column
This gets a table with one row for every year.
```{r}
babynames %>%
  group_by (year) %>%
  summarize(year_total = sum(number))
```
We want to get the total number of people born in each year. For that, change the summarize to a mutate.

=======================================================================================

Combining group_by() and mutate()
```{r}
babnames %>%
  group_by (year) %>%
  mutate(year_total = sum(number)) #Grouped Mutate
```
Just like group_by() and summarize() work well together
group_by() and mutate are a great pair. 

=======================================================================================

ungroup()

```{r}
babynames %>%
  group_by(year) %>%
  mutate(year_total = sum(number)) %>%
  ungroup() %>%
  mutate(fraction = number / year_total) #the fraction of people born in each year that have each name
```

=======================================================================================

Finding the year each name is most common

We are going to explore which year each name was the most common.
To do this, you'll be combining the grouped mutate approach with a top_n.
```{r}
# Find the year each name is most common 
babynames %>%
  group_by(year) %>%
  mutate(year_total = sum(number)) %>%
  ungroup() %>%
  mutate(fraction = number / year_total) %>%
  group_by(name) %>%
  top_n(1, fraction)
```
# Groups:   name [48,040]
    year name      number year_total  fraction
   <dbl> <chr>      <int>      <int>     <dbl>
 1  1880 Abbott         5     201478 0.0000248
 2  1880 Abe           50     201478 0.000248 
 3  1880 Abner         27     201478 0.000134 
 4  1880 Adelbert      28     201478 0.000139 
 5  1880 Adella        26     201478 0.000129 
 6  1880 Adolf          6     201478 0.0000298
 7  1880 Adolph        93     201478 0.000462 
 8  1880 Agustus        5     201478 0.0000248
 9  1880 Albert      1493     201478 0.00741  
10  1880 Albertina      7     201478 0.0000347

=======================================================================================

Visualizing the normalized change in popularity

We picked a few names and calculated each of them as a fraction of their peak. 

```{r}
# Filter for the names Steven, Thomas, and Matthew
names_filtered <- names_normalized %>%
  filter(name %in% c("Steven", "Thomas", "Matthew"))

# Visualize these names over time
ggplot(names_filtered, aes(x = year, y = fraction_max, color = name)) +
  geom_line()
```

=======================================================================================

Window function

This function takes a vector, and returns another vector of the same length. 
We will be learning to use the lag() function. We can lag a vector, which means moving each item to the right by one.
```{r}
babynames_fraction %>%
  filter(name == "Matthew") %>%
  arrange(year) %>%
  mutate(difference = fraction - lag(fraction))
```
We want to take each fraction, and subtract the "lagged" fraction, with fraction minus lag(fraction).

=======================================================================================

Using ratios to describe the frequency of a name

We've learned how to find the difference in the frequency of a baby name between consecutive years. 
What if instead of finding the difference, we find the ratio?
```{r}
babynames_fraction %>%
  # Arrange the data in order of name, then year 
  arrange(name, year) %>%
  # Group the data by name
  group_by(name) %>%
  # Add a ratio column that contains the ratio between each year 
  mutate(ratio = fraction / lag(fraction))
```
# Groups:   name [48,040]
    year name    number year_total   fraction  ratio
   <dbl> <chr>    <int>      <int>      <dbl>  <dbl>
 1  2010 Aaban        9    3672066 0.00000245 NA    
 2  2015 Aaban       15    3648781 0.00000411  1.68 
 3  1995 Aadam        6    3652750 0.00000164 NA    
 4  2000 Aadam        6    3767293 0.00000159  0.970
 5  2005 Aadam        6    3828460 0.00000157  0.984
 6  2010 Aadam        7    3672066 0.00000191  1.22 
 7  2015 Aadam       22    3648781 0.00000603  3.16 
 8  2010 Aadan       11    3672066 0.00000300 NA    
 9  2015 Aadan       10    3648781 0.00000274  0.915
10  2000 Aadarsh      5    3767293 0.00000133 NA    

======================================================================================= Data Manipulation with dplyr .